---
title: "Data_analysis_PE_rhizoboxes"
author: "Benjamin Delory and Inés Alonso-Crespo"
date: "06/08/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    fig_caption: TRUE
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

## Load R packages

```{r Load R packages, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggpubr)
library(lme4)
library(emmeans)
library(lmerTest)
library(Hmisc)
library(ggeffects)
library(MASS)
library(viridis)
library(sp)
library(stringr)
```

## Create function to calculate MRD and D90 from root biomass data

The `vrd` function calculates the following parameters for each rhizobox:

* **RDWtotal**: the total root dry weight (mg) 
* **MRD**: mean rooting depth (cm) following Mommer et al (2010) [Unveiling below-ground species abundance in a biodiversity experiment: a test of vertical niche differentiation among grassland species](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2745.2010.01702.x). *Journal of Ecology*, 98, 1117-1127. 
* **beta**: parameter of the asymptotic equation used to model the relationship between the cumulative proportion of root biomass (C) and soil depth (d). A description of the model can be found in Gale and Grigal (1987). [Vertical root distributions of northern tree species in relation to successional status](https://www.nrcresearchpress.com/doi/10.1139/x87-131#.Xyp3UjVS-70). *Can J For Res*, 17, 829-834.  
$$C=1-\beta^d$$ 
* **D90**: depth value (in cm) above which 90% of the root biomass lies.
$$D90=\frac{log(0.1)}{log(\beta)}$$

```{r Create vrd function, message=FALSE, warning=FALSE}
vrd<-function(x, data, depth, replicate, D=90){
  
  #x is a formula: response~factor
  #data is a data frame with the raw data
  #depth is the name of the column with maxDepth values (character string)
  #replicate is the name of the column with replicate IDs (character string)
  #D is the depth threshold (in %)
  
  if (class(x)=="formula"){
    
    response<-as.character(x[[2]])
    factor<-as.character(x[[3]])}
  
  res<-aggregate(data[,response], by=list(Treatment=data[,factor], Rz=data[,"Rz"], Replicate=data[,replicate]), sum)
  meanRDW<-aggregate(data[,response], by=list(Treatment=data[,factor], Rz=data[,"Rz"], Replicate=data[,replicate]), mean)
  sdRDW<-aggregate(data[,response], by=list(Treatment=data[,factor], Rz=data[,"Rz"], Replicate=data[,replicate]), sd)
  colnames(res)[4]<-"RDWtotal"
  res$MRD<-NA
  res$beta<-NA
  res<-res[order(res$Treatment),]
  
  for (i in 1:nrow(res)){
    
    sub<-data[data[,factor]==res$Treatment[i] & data[,replicate]==res$Replicate[i], c(factor, replicate, depth, response)]
    sub[,factor]<-as.character(sub[,factor])
    sub[nrow(sub)+1,]<-c(sub[1,factor], sub[1,replicate], 0, 0)
    sub[,response]<-as.numeric(sub[,response])
    sub[,depth]<-as.numeric(sub[,depth])
    sub<-sub[order(sub[,depth]),]
    sub[,response]<-cumsum(sub[,response])/sum(sub[,response])
    
    r<-sub[,response]
    d<-sub[,depth]
    
    model<-nls(r~1-beta^d, start=list(beta=1)) #Fit model
    
    res$beta[i]<-coef(model)
    
    res$MRD[i]<-sum((sub[,response]/sum(sub[,response]))*(sub[,depth]-5))}
  
  if (D<100) {res[,paste("D", D, sep="")]<-log(1-(D/100))/log(res$beta)} else {stop("D must be smaller than 100")}
  
  res$invCV<-meanRDW$x/sdRDW$x
  
  return(res[,c(2,1,3:8)])}
```

## Create function to calculate unstandardised effect sizes

`bootstrap_ES` is an R function calculating unstandardised effect sizes for each treatment and their compatibility intervals (bootstrap resampling using the percentile method) for one response variable. The number of bootstrap resamples is set using `n_perm` (default to 20000). The argument `plot` can be used to create a histogram displaying the results.

This function is a modified version of the [function coded by Masahiro Ryo](https://github.com/masahiroryo/joint-ES-estimate/blob/master/20190801_analysis_v2.0.R) for the following paper: [Rillig et al (2019). The role of multiple global change factors in driving soil functions and microbial biodiversity. Science, 366, 886-890](https://science.sciencemag.org/content/366/6467/886).

```{r bootstrap_ES, message=FALSE, warning=FALSE}
bootstrap_ES <- function(response, data=data, contr, n_perm = 20000, plot=FALSE){
  
  #response is the response variable
  #data is a data frame with the raw data
  #contr is a vector of contrasts (factor level separated by "/", no space)
  
  summary <- data.frame(Contrast=NA, lowCI=NA, ES=NA, highCI=NA, relES=NA)
  alles <- data.frame()
  row<-0
  
    for (c in contr){
      
      split<-strsplit(c, "/")[[1]]
      
      bs <- numeric(0)
      population_1 <- na.omit(data[data$Treatment==split[1], response])
      population_2 <- na.omit(data[data$Treatment==split[2], response])
      size_1 <- length(population_1)
      size_2 <- length(population_2)
      
      if (size_1==0|size_2==0){} else {
        
        row<-row+1
        
        set.seed(123) #For reproducibility
        
        for (i in c(1:n_perm)){
          
          k_1 <- mean(sample(population_1, size_1, replace = T))
          k_2 <- mean(sample(population_2, size_2, replace = T))
          bs <- append(bs, k_1-k_2)}
      
      summary[row,1] <- c
      summary[row,2:4] <- c(quantile(bs, .025), mean(population_1)-mean(population_2), quantile(bs, .975))
      summary[row,5] <- 100*(mean(population_1)-mean(population_2))/mean(population_2)
      alles<-rbind(alles, data.frame(Contrast=rep(c, n_perm), ES=bs))}
  
      if (plot==TRUE){
        hist(bs, main=c, xlab=paste("ES", response, sep=" "))
        points(mean(population_1)-mean(population_2), 0, pch=16, cex=2, col="red")
        segments(x0=quantile(bs, .025), y0=0, x1=quantile(bs, .975), y1=0, lwd=2, col="red")}}
  
  return(list(summary=summary, bootstrap=alles))}
```

## Analysis of shoot biomass data

### Load data

```{r Load data, message=FALSE, warning=FALSE}
shoot <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Data_shoot_biomass_RZ_PE_2017.txt")
shoot<-as.data.frame(shoot)
shoot$Treatment<-as.factor(shoot$Treatment)

#Convert data in mg
shoot[,4:16]<-shoot[,4:16]*1000

#Remove rhizoboxes where Lotus corniculatus did not grow
shoot<-na.omit(shoot)

#Prepare data for ggplot
shoot1<-data.frame(Treatment=rep(shoot$Treatment, 9),
                   Replicate=rep(shoot$Replicate, 9),
                   Species=rep(c("Trifolium pratense", "Lotus corniculatus", "Medicago sativa", "Festuca rubra", "Holcus lanatus", "Dactylis glomerata", "Centaurea jacea", "Leucanthemum vulgare", "Achillea millefolium"), each=31),
                   SDW=as.vector(as.matrix(shoot[,4:12])))
```

### Plot shoot biomass for each species

```{r Plot shoot biomass per species, fig.height=6.69, fig.width=7.09, message=FALSE, warning=FALSE}
shoot1$Treatment<-factor(shoot1$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
shoot1$Species<-factor(shoot1$Species, levels=c("Achillea millefolium", "Centaurea jacea", "Leucanthemum vulgare", "Dactylis glomerata", "Festuca rubra", "Holcus lanatus", "Lotus corniculatus", "Medicago sativa","Trifolium pratense"))
                                                
colors<-c("#999999", "#56B4E9", "#E69F00", "#009E73", "black")

(p1<-ggplot(shoot1, aes(x=Treatment, y=SDW, color=Treatment))+
      facet_wrap(~Species, scales="free_y")+
      geom_jitter(shape=1, width=0.15)+
      stat_summary(fun.data="mean_cl_boot", size=0.5)+
      theme_bw()+
      theme(legend.position = "none")+
      ylab("Shoot dry weight (mg)")+
      xlab("")+
      theme(axis.text=element_text(size=10, color="black"),
            strip.text=element_text(face="italic", size=10),
            axis.title.y=element_text(margin=margin(r=10)),
            axis.text.x = element_text(angle=90, vjust=0.5))+
      scale_color_manual(values=colors))

#ggsave("Figure3.tiff", p1, dpi=1000, compression="lzw", width=18, height=17, units="cm")
```

### Plot total shoot biomass

```{r Plot total shoot biomass, message=FALSE, warning=FALSE, fig.height=3.54, fig.width=3.54}
shoot$Treatment<-factor(shoot$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
colors<-c("#999999", "#56B4E9", "#E69F00", "#009E73", "black")

(p2<-ggplot(shoot, aes(x=Treatment, y=Total, color=Treatment)) + 
    geom_jitter(shape=1, width=0.15) + 
    labs(x = "", y = "Total shoot dry weight (mg)") +
    theme_bw()+
    stat_summary(fun.data="mean_cl_boot", size=0.5)+
    scale_color_manual(values=colors)+
    theme(legend.position = "none",
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
          axis.text.x=element_text(angle=0)))

#ggsave("Total_shoot_productivity.tiff", p2, dpi=1000, compression="lzw", width=9, height=9, units="cm")
```

### Effect sizes

#### *Achillea millefolium*

```{r, fig.width=4, fig.height=4}
es.Am<-bootstrap_ES(response="Am", data=shoot, contr=c("F-first/Sync1", "G-first/Sync2", "L-first/Sync2", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Am$summary)

ggplot(es.Am$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Am$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Am$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Centaurea jacea*

```{r, fig.width=4, fig.height=4}
es.Cj<-bootstrap_ES(response="Cj", data=shoot, contr=c("F-first/Sync1", "G-first/Sync2", "L-first/Sync2", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Cj$summary)

ggplot(es.Cj$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Cj$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Cj$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Leucanthemum vulgare*

```{r, fig.width=4, fig.height=4}
es.Lv<-bootstrap_ES(response="Lv", data=shoot, contr=c("F-first/Sync1", "G-first/Sync2", "L-first/Sync2", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Lv$summary)

ggplot(es.Lv$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Lv$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Lv$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Dactylis glomerata*

```{r, fig.width=4, fig.height=4}
es.Dg<-bootstrap_ES(response="Dg", data=shoot, contr=c("G-first/Sync1", "F-first/Sync2", "L-first/Sync2", "F-first/L-first"), plot=FALSE)
knitr::kable(es.Dg$summary)

ggplot(es.Dg$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Dg$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Dg$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Festuca rubra*

```{r, fig.width=4, fig.height=4}
es.Fr<-bootstrap_ES(response="Fr", data=shoot, contr=c("G-first/Sync1", "F-first/Sync2", "L-first/Sync2", "F-first/L-first"), plot=FALSE)
knitr::kable(es.Fr$summary)

ggplot(es.Fr$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Fr$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Fr$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Holcus lanatus*

```{r, fig.width=4, fig.height=4}
es.Hl<-bootstrap_ES(response="Hl", data=shoot, contr=c("G-first/Sync1", "F-first/Sync2", "L-first/Sync2", "F-first/L-first"), plot=FALSE)
knitr::kable(es.Hl$summary)

ggplot(es.Hl$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Hl$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Hl$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Lotus corniculatus*

```{r, fig.width=4, fig.height=4}
es.Lc<-bootstrap_ES(response="Lc", data=shoot, contr=c("L-first/Sync1", "F-first/Sync2", "G-first/Sync2", "F-first/G-first"), plot=FALSE)
knitr::kable(es.Lc$summary)

ggplot(es.Lc$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Lc$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Lc$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Medicago sativa*

```{r, fig.width=4, fig.height=4}
es.Ms<-bootstrap_ES(response="Ms", data=shoot, contr=c("L-first/Sync1", "F-first/Sync2", "G-first/Sync2", "F-first/G-first"), plot=FALSE)
knitr::kable(es.Ms$summary)

ggplot(es.Ms$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Ms$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Ms$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Trifolium pratense*

```{r, fig.width=4, fig.height=4}
es.Tp<-bootstrap_ES(response="Tp", data=shoot, contr=c("L-first/Sync1", "F-first/Sync2", "G-first/Sync2", "G-first/F-first"), plot=FALSE)
knitr::kable(es.Tp$summary)

ggplot(es.Tp$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Tp$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Tp$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### Total shoot biomass

```{r, fig.width=4, fig.height=4}
es.Total<-bootstrap_ES(response="Total", data=shoot, contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Total$summary)

ggplot(es.Total$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Total$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Total$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

## Analysis of root biomass data

### Load root biomass data

```{r Load root biomass data}
root <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Data_root_biomass_RZ_PE_2017.txt")
root<-as.data.frame(root)
root$Treatment<-as.factor(root$Treatment)

root$RDW<-root$RDW*1000 #Convert to mg
root<-root[-which(root$Rz==137),] #Root biomass in layer 5 much greater than in layers 4 and 3 (this might be due to a data encoding mistake)
```

### Plot total root productivity

```{r Plot total root productivity, message=FALSE, warning=FALSE, fig.height=3.54, fig.width=3.54}
root$Treatment<-factor(root$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
colors<-c("#999999", "#56B4E9", "#E69F00", "#009E73", "black")

root1<-vrd(x=RDW~Treatment, data=root, depth="maxDepth", replicate="Replicate", D=90)

(p3<-ggplot(root1, aes(x=Treatment, y=RDWtotal, color=Treatment)) + 
      geom_jitter(shape=1, width=0.1) + 
      labs(x = "", y = "Total root dry weight (mg)") +
      theme_bw()+
      stat_summary(fun.data="mean_cl_boot", size=0.5)+
      scale_color_manual(values=colors)+
      theme(legend.position = "none",
            axis.text=element_text(size=10, color="black"),
            axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
            axis.text.x=element_text(angle=0))+
      scale_y_continuous(limits=c(0,300), breaks=c(0,50,100,150,200,250,300)))

(p<-ggplot(root1, aes(x=Treatment, y=invCV, color=Treatment)) + 
      geom_jitter(shape=1, width=0.1) + 
      labs(x = "", y = "Community inverse CV") +
      theme_bw()+
      stat_summary(fun.data="mean_cl_boot", size=0.5)+
      scale_color_manual(values=colors)+
      theme(legend.position = "none",
            axis.text=element_text(size=10, color="black"),
            axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
            axis.text.x=element_text(angle=0)))

#ggsave("Total_root_productivity.tiff", p3, dpi=1000, compression="lzw", width=9, height=9, units="cm")
```

### Effect sizes

#### Total root productivity

```{r, fig.width=4, fig.height=4}
es.rdw<-bootstrap_ES(response="RDWtotal", data=root1, contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.rdw$summary)

ggplot(es.rdw$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.rdw$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.rdw$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES RDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### MRD biomass

```{r, fig.width=4, fig.height=4}
es.mrdb<-bootstrap_ES(response="MRD", data=root1, contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.mrdb$summary)

ggplot(es.mrdb$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.mrdb$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.mrdb$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES MRD (cm)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

### Create combined plot for aboveground and belowground productivity (figure 2)

```{r Create figure 2, message=FALSE, warning=FALSE, fig.height=3.54, fig.width=7.08}
(figure2<-ggarrange(p2, p3, align="hv", nrow=1, ncol=2, labels=c("(a)", "(b)"), font.label=list(size=12)))
#ggsave("Figure2.tiff", figure2, dpi=1000, compression="lzw", width=18, height=9, units="cm")
```

### Plot root biomass distribution in soil layers

```{r Plot root biomass distribution, message=FALSE, warning=FALSE, fig.width=7.09, fig.height=5.12}

supp.labs<-c("0-10 cm", "10-20 cm", "20-30 cm", "30-40 cm", "40-50 cm", ">50 cm")
names(supp.labs)<-c("0-10 cm", "10-20 cm", "20-30 cm", "30-40 cm", "40-50 cm", "50-60 cm")
  
(p4<-ggplot(root, aes(x=Treatment, y=RDW, color=Treatment)) + 
    geom_jitter(shape=1, width=0.1) + 
    labs(x = "", y = "Total root dry weight (mg)") +
    theme_bw()+
    stat_summary(fun.data="mean_cl_boot", size=0.5)+
    scale_color_manual(values=colors)+
    theme(legend.position = "none",
          axis.text=element_text(color="black"),
          axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
          axis.text.x=element_text(size=10, angle=90),
          axis.text.y=element_text(size=10),
          strip.text = element_text(size=10))+
    facet_wrap(~Layer, scales="free_y",
               labeller=labeller(Layer=supp.labs)))

#ggsave("Vertical_root_distribution.tiff", p4, dpi=1000, compression="lzw", width=18, height=13, units="cm")

```

### Root biomass distribution in the soil

Because some layers did not contain any roots (RDW=0), we had zero-inflated continuous data ($Y \ge 0$). Data were analysed using a hurdle model combining a gamma GLMM and a Bernoulli GLMM (zero-altered gamma model or ZAG model). In a ZAG model, zeroes are generated by the binary part of the model (1 or 0). Non-zero data (that occur that roots are present in a layer) are modeled using a second (continuous) distribution (gamma in this case). The hurdle model was fitted according to Chapter 7 in Zuur and Ieno (2016) [Beginner's guide to zero-inflated models with R](http://highstat.com/index.php/beginner-s-guide-to-zero-inflated-models). Highland Statistics Ltd. 

#### Fit gamma component of hurdle model

```{r Fit gamma model, message=FALSE, warning=FALSE}
root$Depth<-root$maxDepth-5

root.pos<-subset(root, RDW>0) #Only select positive RDW values
root.pos$Rz<-factor(root.pos$Rz)

#Fit GLMM with gamma distribution
H1<-glmmPQL(fixed=RDW~Depth*Treatment, random=~1|Rz, data=root.pos, family=Gamma(link="log"))
summary(H1)

#Model validation
E1<-resid(H1, type="pearson")
F1<-fitted(H1, type="response")
plot(F1, E1, xlab="Fitted values", ylab="Residuals"); abline(h=0)
boxplot(E1~root.pos$Treatment, xlab="", ylab="Residuals")
boxplot(E1~root.pos$Depth, xlab="Depth (cm)", ylab="Residuals")

#Make predictions
pred.mm <- data.frame(Depth=rep(seq(5,55,by=0.1), 5), Treatment=rep(unique(root$Treatment), each=501))
pred.mm$Pred <- predict(H1, newdata=pred.mm, type="response", level=0)

#Plot GLMM fit
(p5<-ggplot(pred.mm) +
    geom_jitter(data=root.pos, aes(x=Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(aes(x = Depth, y = Pred, color=Treatment), size=1) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))

#Plot GLMM fit
#One curve for each rhizobox
(p6<-ggplot(root.pos) +
    geom_jitter(aes(x=Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(data=cbind(root.pos, Pred=predict(H1, type="response")), aes(x = Depth, y = Pred, color=Treatment, group=Rz), size=1) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))
```
 
#### Fit binomial component of hurdle model

```{r Fit binomial model, message=FALSE, warning=FALSE}
root$RDW.01<-as.numeric(root$RDW > 0)
root$Rz<-factor(root$Rz)
root$sDepth<-scale(root$Depth, center=TRUE, scale=TRUE)

#Fit GLMM with a binomial distribution
H2<-glmmPQL(fixed=RDW.01~Depth*Treatment, random=~1|Rz, data=root, family=binomial(link="logit"))
summary(H2)

#Make predictions
pred.mm <- data.frame(Depth=rep(seq(5,55,by=0.1), 5),
                      Treatment=rep(unique(root$Treatment), each=501))
pred.mm$Pred <- predict(H2, newdata=pred.mm, type="response", level=0)

#Plot GLMM fit
(p7<-ggplot(pred.mm) +
    geom_jitter(data=root, aes(x=Depth, y=RDW.01, colour=Treatment), height=0.2, width=1, shape=1) +
    geom_line(aes(x = Depth, y = Pred, color=Treatment), size=1) +
    ylab("Probability that roots are present")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))
```

#### Merge gamma and binomial components

```{r Create hurdle model, message=FALSE, warning=FALSE, fig.height=3.94, fig.width=3.94}
beta<-fixef(H1) #Extract regression coefficients of Gamma GLMM
gamma<-fixef(H2) #Extract regression coefficients of Binomial GLMM

#Create new data for predictions
newdata <- expand.grid(Depth = seq(5, 55, by = 0.1),
                       Treatment = levels(root$Treatment))

X<-model.matrix(~Depth*Treatment, data=newdata)

mu <- exp(X %*% beta)
Pi  <- exp(X %*% gamma) / (1 + exp(X %*% gamma))

#Get predictions for Hurdle model
newdata$PredZAG<-Pi * mu

#Plot hurdle model fit
(p8<-ggplot(root) +
    geom_jitter(data=root, aes(x=Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(data=newdata, aes(x = Depth, y = PredZAG, color=Treatment), size=0.8) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          legend.position=c(0.85,0.75),
          legend.text=element_text(size=8),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))

#ggsave("Figure4.tiff", p8, dpi=1000, compression="lzw", width=10, height=9.5, units="cm")
```

## Analysis of root distribution over time

### Load RootPainter data

```{r Load RootPainter data, message=FALSE, warning=FALSE}
root.depth <- read.csv("Z:/Poem/Rhizoboxes/Backup_RootPainter/drive_rp_sync/projects/PE_rhizotrons_2017_1/results/segmentations/results.csv")

ID <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Image_RZ.txt")
TIME <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Date_TimePoint.txt")

filenames<-sapply(str_split(as.character(root.depth$image), pattern="-"), function(x){return(x[5])})
filenames<-as.numeric(sapply(str_split(filenames, pattern="\\."), function(x){return(x[1])}))

root.depth$Date<-sapply(str_split(as.character(root.depth$image), pattern="-"), function(x){
  return(paste(x[1], x[2], x[3], sep="-"))})

root.depth$imageID<-filenames
root.depth$Rz<-ID$RZ[match(filenames, ID$Image)]
root.depth$Time<-TIME$Day[match(root.depth$Date, TIME$Date)]

root.depth<-root.depth[order(root.depth$Rz, root.depth$Time),]

#Add treatment column
root.depth$Treatment<-root1$Treatment[match(root.depth$Rz, root1$Rz)]

#Set Sync2 to zero before the second sowing
root.depth$mrd[which(root.depth$Treatment=="Sync2" & root.depth$Time<=9)]<-0

#Remove rhizoboxes 35, 70, 15, 27, 137
root.depth<-root.depth[-which(root.depth$Rz==35|root.depth$Rz==70|root.depth$Rz==15|root.depth$Rz==27|root.depth$Rz==137),]

root.depth$Treatment<-factor(root.depth$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
```

### Plot mean rooting depth over time

```{r Calculate MRD, message=FALSE, warning=FALSE}
#Create empty dataset for summary statistics
stat.mrd<-data.frame(Treatment=rep(NA, length(unique(root.depth$Rz))),
                     Time=rep(NA, length(unique(root.depth$Rz))),
                     meanValue=rep(NA, length(unique(root.depth$Rz))),
                     lower=rep(NA, length(unique(root.depth$Rz))),
                     upper=rep(NA, length(unique(root.depth$Rz))))
 
k<-0
 
for (i in unique(root.depth$Treatment)){
   
  for (j in unique(root.depth$Time)){
     
    k<-k+1
    sub<-root.depth[root.depth$Treatment==i & root.depth$Time==j,]
    #Calculate compatibility interval (bootstrap resampling)
    set.seed(1)
    a<-smean.cl.boot(sub$mrd, B=10000)
    stat.mrd[k,]<-c(i, j, as.numeric(a))}}
 
stat.mrd[,2]<-as.numeric(stat.mrd[,2])
stat.mrd[,3]<-as.numeric(stat.mrd[,3])
stat.mrd[,4]<-as.numeric(stat.mrd[,4])
stat.mrd[,5]<-as.numeric(stat.mrd[,5])
 
stat.mrd$Treatment<-factor(stat.mrd$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))

#Plot mean rooting depth over time
(p9<-ggplot(stat.mrd, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MRD (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(stat.mrd$Time)))+
  scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1))

#Plot mean rooting depth over time
#One line per rhizobox
(p10<-ggplot(root.depth, aes(x=Time, y=mrd, color=Treatment, group=Rz))+
  geom_line()+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MRD (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors))

#Plot mean rooting depth at last observation date

(p11<-ggplot(root.depth[root.depth$Time==44,], aes(x=Treatment, y=mrd, colour=Treatment)) +
    geom_jitter(width=0.1, shape=1) + 
    stat_summary(fun.data="mean_cl_boot")+
    ylab("MRD (cm)")+
    xlab("")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10))))

model<-glm(-mrd~Treatment, root.depth[root.depth$Time==44,], family=Gamma(link="log"))

summary(model)
summary(emmeans(model, "Treatment", contr="tukey"), type="response")
summary(emmeans(model, "Treatment", contr="tukey"))
```

### Fit linear mixed effect model

```{r Fit lmer to mean rooting depth data, message=FALSE, warning=FALSE}
#Calculate number of days after first sowing (dafs)
root.depth$dafs<-root.depth$Time
root.depth$dafs[root.depth$Treatment=="Sync2"]<-root.depth$dafs[root.depth$Treatment=="Sync2"]-10
root.depth1<-root.depth[root.depth$dafs>=0,]

#Fit linear mixed-effect model(random slope and random intercept)
model<-lmer(mrd~Treatment*dafs+(dafs|Rz), root.depth1)
summary(model)
plot(model)

#Plot model fit
#One line per rhizobox
(p12<-ggplot(root.depth1, aes(x=dafs, y=mrd, color=Treatment, group=Rz))+
  geom_line(data = cbind(root.depth1, pred = predict(model)), aes(y = pred), size = 1)+
  geom_point(shape=16)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MRD (cm)")+
  xlab("Days after sowing")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(limits=c(1, 44)))

pred.mm <- ggpredict(model, terms = c("dafs [1,2,4,6,7,8,9,11,13,14,15,16,18,20,21,22,23,25,27,28,29,30,32,34,35,37,39,42,44]", "Treatment"))

data.mm<-data.frame(dafs=pred.mm$x,
                    Pred=pred.mm$predicted,
                    Treatment=pred.mm$group,
                    CIlow=pred.mm$conf.low,
                    CIhigh=pred.mm$conf.high)

data.mm<-data.mm[-which(data.mm$Treatment=="Sync2" & data.mm$dafs>34),]
data.mm<-data.mm[-which(data.mm$Treatment!="Sync2" & data.mm$dafs<2),]

#Plot model fit
(p13<-ggplot(data.mm) + 
    geom_line(aes(x = dafs, y = Pred, group=Treatment, color=Treatment), size=1) +   
    geom_ribbon(aes(x = dafs, ymin = CIlow, ymax = CIhigh, fill=Treatment, alpha=Treatment)) +
    ylab("MRD (cm)")+
    xlab("Days after sowing")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    scale_fill_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1)))
```

### Correlation between MRD biomass and MRD images

```{r Correlation MRD, message=FALSE, warning=FALSE}
root1<-root1[order(root1$Rz),] #Contain MRD values calculated from root biomass
root1$MRDrp<-root.depth[root.depth$Time==44,"mrd"]

corrMRD<-data.frame(MRDb=root1$MRD,
                    MRDi=root1$MRDrp,
                    Treatment=root1$Treatment)

(p14<-ggplot(corrMRD, aes(x=-MRDi, y=MRDb, colour=Treatment))+
      geom_smooth(method="lm", colour="black")+
      geom_point(shape=16)+
      theme_bw()+
      theme(legend.title = element_blank(),
            axis.text=element_text(size=10, color="black"),
            axis.title.y = element_text(margin=margin(r=10)),
            axis.title.x = element_text(margin=margin(t=10)))+
      scale_color_manual(values=colors)+
      xlab("MRD - image analysis (cm)")+
      ylab("MRD - biomass (cm)"))
```

### Create figure MRD over time

```{r Figure MRD over time, message=FALSE, warning=FALSE, fig.height=2.36, fig.width=7.09}
text.size<-8
title.size<-9

p14<-ggplot(corrMRD, aes(x=-MRDi, y=MRDb, colour=Treatment))+
      geom_smooth(method="lm", colour="black", size=1)+
      geom_point(shape=16)+
      theme_bw()+
      theme(legend.title=element_blank(),
            legend.position=c(0.85,0.24),
            legend.text=element_text(size=7),
            legend.key.size = unit(0.7, "line"),
            legend.background = element_blank(),
            axis.text=element_text(size=text.size, color="black"),
            axis.title=element_text(size=title.size),
            axis.title.y = element_text(margin=margin(r=5)),
            axis.title.x = element_text(margin=margin(t=5)))+
      scale_color_manual(values=colors)+
      xlab("MRD - image analysis (cm)")+
      ylab("MRD - biomass (cm)")

p9<-ggplot(stat.mrd, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16, size=1)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        legend.position=c(0.85,0.82),
        legend.text=element_text(size=7),
        legend.key.size = unit(0.7, "line"),
        legend.background = element_blank(),
        axis.text=element_text(size=text.size, color="black"),
        axis.title=element_text(size=title.size),
        axis.title.y = element_text(margin=margin(r=5)),
        axis.title.x = element_text(margin=margin(t=5)),
        panel.grid.minor = element_blank())+
  ylab("Mean rooting depth (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(root.depth$Time),
                     labels=c(2,4,"",9,"","",16,"","",23,"","",30,"","",37, "","",44))+
  scale_alpha_manual(values=c(0.2,0.2,0.2,0.2,0.1))

p13<-ggplot(data.mm) + 
    geom_line(aes(x = dafs, y = Pred, group=Treatment, color=Treatment), size=0.6) +
    geom_ribbon(aes(x = dafs, ymin = CIlow, ymax = CIhigh, fill=Treatment, alpha=Treatment), color=NA) +
    ylab("Mean rooting depth (cm)")+
    xlab("Days after sowing")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    scale_fill_manual(values=colors)+
    theme(legend.position="none",
          axis.text=element_text(size=text.size, color="black"),
          axis.title=element_text(size=title.size),
          axis.title.y = element_text(margin=margin(r=5)),
          axis.title.x = element_text(margin=margin(t=5)),
          panel.grid.minor = element_blank())+
    scale_alpha_manual(values=c(0.2,0.2,0.2,0.2,0.1))+
    scale_x_continuous(breaks=seq(0,44, by=4),
                     labels=seq(0,44, by=4))+
    scale_y_continuous(breaks=c(0,-4,-8,-12,-16), labels=c(0,-4,-8,-12,-16), limits=c(-16,0))

(p15<-ggarrange(p14, p9, p13, nrow=1, ncol=3, labels=c("(a)", "(b)", "(c)"), font.label = list(size=11)))

#ggsave("Figure5.tiff", p15, dpi=1000, compression="lzw", width=20, height=6, units="cm")
```