---
title: "Data_analysis_PE_rhizoboxes"
author: "Benjamin Delory and Inés Alonso-Crespo"
date: "15/11/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    fig_caption: TRUE
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

## Load R packages

```{r Load R packages, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggpubr)
library(ggsignif)
library(lme4)
library(emmeans)
library(lmerTest)
library(Hmisc)
library(ggeffects)
library(MASS)
library(viridis)
library(sp)
library(stringr)
library(dplyr)
```

## Create function to calculate MRD and D90 from root biomass data

The `vrd` function calculates the following parameters for each rhizobox:

* **RDWtotal**: the total root dry weight (mg) 
* **MRD**: mean rooting depth (cm) following Mommer et al (2010) [Unveiling below-ground species abundance in a biodiversity experiment: a test of vertical niche differentiation among grassland species](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2745.2010.01702.x). *Journal of Ecology*, 98, 1117-1127. 
* **beta**: parameter of the asymptotic equation used to model the relationship between the cumulative proportion of root biomass (C) and soil depth (d). A description of the model can be found in Gale and Grigal (1987). [Vertical root distributions of northern tree species in relation to successional status](https://www.nrcresearchpress.com/doi/10.1139/x87-131#.Xyp3UjVS-70). *Can J For Res*, 17, 829-834.  
$$C=1-\beta^d$$ 
* **D90**: depth value (in cm) above which 90% of the root biomass lies.
$$D90=\frac{log(0.1)}{log(\beta)}$$

```{r Create vrd function, message=FALSE, warning=FALSE}
vrd<-function(x, data, depth, replicate, D=90){
  
  #x is a formula: response~factor
  #data is a data frame with the raw data
  #depth is the name of the column with maxDepth values (character string)
  #replicate is the name of the column with replicate IDs (character string)
  #D is the depth threshold (in %)
  
  if (class(x)=="formula"){
    
    response<-as.character(x[[2]])
    factor<-as.character(x[[3]])}
  
  res<-aggregate(data[,response], by=list(Treatment=data[,factor], Rz=data[,"Rz"], Replicate=data[,replicate]), sum)
  meanRDW<-aggregate(data[,response], by=list(Treatment=data[,factor], Rz=data[,"Rz"], Replicate=data[,replicate]), mean)
  sdRDW<-aggregate(data[,response], by=list(Treatment=data[,factor], Rz=data[,"Rz"], Replicate=data[,replicate]), sd)
  colnames(res)[4]<-"RDWtotal"
  res$MRD<-NA
  res$beta<-NA
  res<-res[order(res$Treatment),]
  
  for (i in 1:nrow(res)){
    
    sub<-data[data[,factor]==res$Treatment[i] & data[,replicate]==res$Replicate[i], c(factor, replicate, depth, response)]
    sub[,factor]<-as.character(sub[,factor])
    sub[nrow(sub)+1,]<-c(sub[1,factor], sub[1,replicate], 0, 0)
    sub[,response]<-as.numeric(sub[,response])
    sub[,depth]<-as.numeric(sub[,depth])
    sub<-sub[order(sub[,depth]),]
    sub[,response]<-cumsum(sub[,response])/sum(sub[,response])
    
    r<-sub[,response]
    d<-sub[,depth]
    
    model<-nls(r~1-beta^d, start=list(beta=1)) #Fit model
    
    res$beta[i]<-coef(model)
    
    res$MRD[i]<-sum((sub[,response]/sum(sub[,response]))*(sub[,depth]-5))}
  
  if (D<100) {res[,paste("D", D, sep="")]<-log(1-(D/100))/log(res$beta)} else {stop("D must be smaller than 100")}
  
  res$invCV<-meanRDW$x/sdRDW$x
  
  return(res[,c(2,1,3:8)])}
```

## Create function to calculate unstandardised effect sizes

`bootstrap_ES` is an R function calculating unstandardised effect sizes for each treatment and their compatibility intervals (bootstrap resampling using the percentile method) for one response variable. The number of bootstrap resamples is set using `n_perm` (default to 20000). The argument `plot` can be used to create a histogram displaying the results.

This function is a modified version of the [function coded by Masahiro Ryo](https://github.com/masahiroryo/joint-ES-estimate/blob/master/20190801_analysis_v2.0.R) for the following paper: [Rillig et al (2019). The role of multiple global change factors in driving soil functions and microbial biodiversity. Science, 366, 886-890](https://science.sciencemag.org/content/366/6467/886).

```{r bootstrap_ES, message=FALSE, warning=FALSE}
bootstrap_ES <- function(response, data=data, contr, n_perm = 20000, plot=FALSE){
  
  #response is the response variable
  #data is a data frame with the raw data
  #contr is a vector of contrasts (factor level separated by "/", no space)
  
  summary <- data.frame(Contrast=NA, lowCI=NA, ES=NA, highCI=NA, relES=NA)
  alles <- data.frame()
  row<-0
  
    for (c in contr){
      
      split<-strsplit(c, "/")[[1]]
      
      bs <- numeric(0)
      population_1 <- na.omit(data[data$Treatment==split[1], response])
      population_2 <- na.omit(data[data$Treatment==split[2], response])
      size_1 <- length(population_1)
      size_2 <- length(population_2)
      
      if (size_1==0|size_2==0){} else {
        
        row<-row+1
        
        set.seed(123) #For reproducibility
        
        for (i in c(1:n_perm)){
          
          k_1 <- mean(sample(population_1, size_1, replace = T))
          k_2 <- mean(sample(population_2, size_2, replace = T))
          bs <- append(bs, k_1-k_2)}
      
      summary[row,1] <- c
      summary[row,2:4] <- c(quantile(bs, .025), mean(population_1)-mean(population_2), quantile(bs, .975))
      summary[row,5] <- 100*(mean(population_1)-mean(population_2))/mean(population_2)
      alles<-rbind(alles, data.frame(Contrast=rep(c, n_perm), ES=bs))}
  
      if (plot==TRUE){
        hist(bs, main=c, xlab=paste("ES", response, sep=" "))
        points(mean(population_1)-mean(population_2), 0, pch=16, cex=2, col="red")
        segments(x0=quantile(bs, .025), y0=0, x1=quantile(bs, .975), y1=0, lwd=2, col="red")}}
  
  return(list(summary=summary, bootstrap=alles))}
```

## Analysis of shoot biomass data

### Load data

```{r Load data, message=FALSE, warning=FALSE}
shoot <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Data_shoot_biomass_RZ_PE_2017.txt")
shoot<-as.data.frame(shoot)
shoot$Treatment<-as.factor(shoot$Treatment)

#Convert data in mg
shoot[,4:16]<-shoot[,4:16]*1000

#Remove rhizoboxes where Lotus corniculatus did not grow
shoot<-na.omit(shoot)

#Prepare data for ggplot
shoot1<-data.frame(Treatment=rep(shoot$Treatment, 9),
                   Replicate=rep(shoot$Replicate, 9),
                   Species=rep(c("Trifolium pratense", "Lotus corniculatus", "Medicago sativa", "Festuca rubra", "Holcus lanatus", "Dactylis glomerata", "Centaurea jacea", "Leucanthemum vulgare", "Achillea millefolium"), each=31),
                   SDW=as.vector(as.matrix(shoot[,4:12])))
```

### Plot total shoot biomass

```{r Plot total shoot biomass, message=FALSE, warning=FALSE, fig.height=3.54, fig.width=3.54}
shoot$Treatment<-factor(shoot$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
colors<-c("#999999", "#56B4E9", "#E69F00", "#009E73", "black")

(p2<-ggplot(shoot, aes(x=Treatment, y=Total, color=Treatment)) + 
    geom_jitter(shape=1, width=0.15) + 
    labs(x = "", y = "Total shoot dry weight (mg)") +
    theme_bw()+
    stat_summary(fun.data="mean_cl_boot", size=0.5)+
    scale_color_manual(values=colors)+
    theme(legend.position = "none",
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
          axis.text.x=element_text(angle=0)))

#ggsave("Total_shoot_productivity.tiff", p2, dpi=1000, compression="lzw", width=9, height=9, units="cm")
```

### Effect sizes

#### *Achillea millefolium*

```{r, fig.width=4, fig.height=4}
es.Am<-bootstrap_ES(response="Am", data=shoot, contr=c("F-first/Sync1", "G-first/Sync2", "L-first/Sync2", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Am$summary)

ggplot(es.Am$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Am$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Am$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Centaurea jacea*

```{r, fig.width=4, fig.height=4}
es.Cj<-bootstrap_ES(response="Cj", data=shoot, contr=c("F-first/Sync1", "G-first/Sync2", "L-first/Sync2", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Cj$summary)

ggplot(es.Cj$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Cj$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Cj$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Leucanthemum vulgare*

```{r, fig.width=4, fig.height=4}
es.Lv<-bootstrap_ES(response="Lv", data=shoot, contr=c("F-first/Sync1", "G-first/Sync2", "L-first/Sync2", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Lv$summary)

ggplot(es.Lv$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Lv$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Lv$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Dactylis glomerata*

```{r, fig.width=4, fig.height=4}
es.Dg<-bootstrap_ES(response="Dg", data=shoot, contr=c("G-first/Sync1", "F-first/Sync2", "L-first/Sync2", "F-first/L-first"), plot=FALSE)
knitr::kable(es.Dg$summary)

ggplot(es.Dg$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Dg$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Dg$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Festuca rubra*

```{r, fig.width=4, fig.height=4}
es.Fr<-bootstrap_ES(response="Fr", data=shoot, contr=c("G-first/Sync1", "F-first/Sync2", "L-first/Sync2", "F-first/L-first"), plot=FALSE)
knitr::kable(es.Fr$summary)

ggplot(es.Fr$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Fr$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Fr$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Holcus lanatus*

```{r, fig.width=4, fig.height=4}
es.Hl<-bootstrap_ES(response="Hl", data=shoot, contr=c("G-first/Sync1", "F-first/Sync2", "L-first/Sync2", "F-first/L-first"), plot=FALSE)
knitr::kable(es.Hl$summary)

ggplot(es.Hl$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Hl$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Hl$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Lotus corniculatus*

```{r, fig.width=4, fig.height=4}
es.Lc<-bootstrap_ES(response="Lc", data=shoot, contr=c("L-first/Sync1", "F-first/Sync2", "G-first/Sync2", "F-first/G-first"), plot=FALSE)
knitr::kable(es.Lc$summary)

ggplot(es.Lc$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Lc$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Lc$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Medicago sativa*

```{r, fig.width=4, fig.height=4}
es.Ms<-bootstrap_ES(response="Ms", data=shoot, contr=c("L-first/Sync1", "F-first/Sync2", "G-first/Sync2", "F-first/G-first"), plot=FALSE)
knitr::kable(es.Ms$summary)

ggplot(es.Ms$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Ms$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Ms$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### *Trifolium pratense*

```{r, fig.width=4, fig.height=4}
es.Tp<-bootstrap_ES(response="Tp", data=shoot, contr=c("L-first/Sync1", "F-first/Sync2", "G-first/Sync2", "G-first/F-first"), plot=FALSE)
knitr::kable(es.Tp$summary)

ggplot(es.Tp$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Tp$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Tp$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### Total shoot biomass

```{r, fig.width=4, fig.height=4}
es.Total<-bootstrap_ES(response="Total", data=shoot, contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.Total$summary)

ggplot(es.Total$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.Total$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.Total$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES SDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

### Plot shoot biomass for each species

```{r Plot shoot biomass per species, fig.height=6.69, fig.width=7.09, message=FALSE, warning=FALSE}
shoot1$Treatment<-factor(shoot1$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
shoot1$Species<-factor(shoot1$Species, levels=c("Achillea millefolium", "Centaurea jacea", "Leucanthemum vulgare", "Dactylis glomerata", "Festuca rubra", "Holcus lanatus", "Lotus corniculatus", "Medicago sativa","Trifolium pratense"))
                                                
colors<-c("#999999", "#56B4E9", "#E69F00", "#009E73", "black")

shoot1$PFG<-"Legumes"
shoot1$PFG[which(shoot1$Species=="Achillea millefolium"|
                   shoot1$Species=="Centaurea jacea"|
                   shoot1$Species=="Leucanthemum vulgare")]<-"Forbs"
shoot1$PFG[which(shoot1$Species=="Dactylis glomerata"|
                   shoot1$Species=="Festuca rubra"|
                   shoot1$Species=="Holcus lanatus")]<-"Grasses"

shoot1$PFG<-factor(shoot1$PFG, levels=c("Forbs", "Grasses", "Legumes"))

data_signif_1<-data.frame(
  Species=c("Achillea millefolium", "Centaurea jacea", "Leucanthemum vulgare", "Dactylis glomerata", "Festuca rubra", "Holcus lanatus", "Lotus corniculatus", "Medicago sativa","Trifolium pratense"),
  PFG=c("Forbs", "Forbs", "Forbs", "Grasses", "Grasses", "Grasses","Legumes","Legumes","Legumes"),
  xmin=c(1,1,1,1,1,1,1,1,1),
  xmax=c(2,2,2,3,3,3,4,4,4),
  x=c(1.5,1.5,1.5,2,2,2,2.5,2.5,2.5),
  y=c(220,255,175,210,65,280,85,360,180),
  annotations=c("*", "*","","","","","","",""),
  linetype=c("y","y",rep("n", 7)))

data_signif_2<-data.frame(
  Species=c("Achillea millefolium", "Achillea millefolium",
            "Centaurea jacea", "Centaurea jacea",
            "Leucanthemum vulgare", "Leucanthemum vulgare",
            "Dactylis glomerata", "Dactylis glomerata",
            "Festuca rubra", "Festuca rubra",
            "Holcus lanatus", "Holcus lanatus",
            "Lotus corniculatus", "Lotus corniculatus",
            "Medicago sativa", "Medicago sativa",
            "Trifolium pratense", "Trifolium pratense"),
  PFG=c("Forbs", "Forbs",
            "Forbs", "Forbs",
            "Forbs", "Forbs",
            "Grasses", "Grasses",
            "Grasses", "Grasses",
            "Grasses", "Grasses",
            "Legumes", "Legumes",
            "Legumes", "Legumes",
            "Legumes", "Legumes"),
  xmin=c(3,4,
         3,4,
         3,4,
         2,4,
         2,4,
         2,4,
         2,3,
         2,3,
         3,2),
  xmax=c(4,5,
         4,5,
         4,5,
         4,5,
         4,5,
         4,5,
         3,5,
         3,5,
         5,5),
  x=c(rep(4,16), 4, 4),
  y=c(40,40,
      190,190,
      50,50,
      60,60,
      20,20,
      50,50,
      28,28,
      130,130,
      65,85),
  annotations=c(rep("",16), "*",""),
  linetype=c(rep("n", 16), "y", "n"))

data_signif_3<-data.frame(
  Species="Trifolium pratense",
  PFG="Legumes",
  xmin=2,
  xmax=3,
  x=2.5,
  y=75,
  annotations="*",
  linetype="y")


data_signif_1$Species<-factor(data_signif_1$Species, levels=c("Achillea millefolium", "Centaurea jacea", "Leucanthemum vulgare", "Dactylis glomerata", "Festuca rubra", "Holcus lanatus", "Lotus corniculatus", "Medicago sativa","Trifolium pratense"))

data_signif_2$Species<-factor(data_signif_2$Species, levels=c("Achillea millefolium", "Centaurea jacea", "Leucanthemum vulgare", "Dactylis glomerata", "Festuca rubra", "Holcus lanatus", "Lotus corniculatus", "Medicago sativa","Trifolium pratense"))

data_signif_3$Species<-factor(data_signif_3$Species, levels=c("Trifolium pratense"))

data_signif_1$PFG<-factor(data_signif_1$PFG, levels=c("Forbs", "Grasses", "Legumes"))

data_signif_2$PFG<-factor(data_signif_2$PFG, levels=c("Forbs", "Grasses", "Legumes"))

data_signif_3$PFG<-factor(data_signif_3$PFG, levels=c("Legumes"))

# New facet label names for species
species.labs <- c("Achillea millefolium (forb)", "Centaurea jacea (forb)", "Leucanthemum vulgare (forb)", "Dactylis glomerata (grass)", "Festuca rubra (grass)", "Holcus lanatus (grass)", "Lotus corniculatus (legume)", "Medicago sativa (legume)","Trifolium pratense (legume)")

names(species.labs) <- c("Achillea millefolium", "Centaurea jacea", "Leucanthemum vulgare", "Dactylis glomerata", "Festuca rubra", "Holcus lanatus", "Lotus corniculatus", "Medicago sativa","Trifolium pratense")

(p1<-ggplot(shoot1)+
      geom_jitter(aes(x=Treatment, y=SDW, color=Treatment), shape=1, width=0.15)+
      stat_summary(aes(x=Treatment, y=SDW, color=Treatment), fun.data="mean_cl_boot", size=0.5)+
      facet_wrap(~Species, scales="free_y", drop=T, labeller=labeller(Species=species.labs))+
      geom_signif(data=data_signif_1,
       mapping=aes(x=x, y_position=y, xmin=xmin, xmax=xmax, annotations=annotations, linetype=linetype),
       manual=T,
       inherit.aes = F,
       vjust=1.4,
       textsize = 6)+
      geom_signif(data=data_signif_2,
       mapping=aes(x=x, y_position=y, xmin=xmin, xmax=xmax, annotations=annotations, linetype=linetype),
       manual=T,
       inherit.aes = F,
       vjust=1.4,
       textsize = 6)+
      geom_signif(data=data_signif_3,
         mapping=aes(x=x, y_position=y, xmin=xmin, xmax=xmax, annotations=annotations, linetype=linetype),
         manual=T,
         inherit.aes = F,
         vjust=1.4,
         textsize = 6)+
      theme_bw()+
      theme(legend.position = "none")+
      ylab("Shoot dry weight (mg)")+
      xlab("")+
      theme(axis.text=element_text(size=10, color="black"),
            strip.text.x =element_text(face="italic", size=9.5),
            axis.title.y=element_text(margin=margin(r=10)),
            axis.text.x = element_text(angle=45, vjust=1, hjust=1))+
      scale_color_manual(values=colors)+
      scale_linetype_manual(values=c(2,1)))

#ggsave("Figure3.tiff", p1, dpi=1000, compression="lzw", width=18, height=17, units="cm")
```

## Analysis of root biomass data

### Load root biomass data

```{r Load root biomass data}
root <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Data_root_biomass_RZ_PE_2017.txt")
root<-as.data.frame(root)
root$Treatment<-as.factor(root$Treatment)

root$RDW<-root$RDW*1000 #Convert to mg
root<-root[-which(root$Rz==137),] #Root biomass in layer 5 much greater than in layers 4 and 3 (this might be due to a data encoding mistake)
```

### Plot total root productivity

```{r Plot total root productivity, message=FALSE, warning=FALSE, fig.height=3.54, fig.width=3.54}
root$Treatment<-factor(root$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))
colors<-c("#999999", "#56B4E9", "#E69F00", "#009E73", "black")

root1<-vrd(x=RDW~Treatment, data=root, depth="maxDepth", replicate="Replicate", D=90)

(p3<-ggplot(root1, aes(x=Treatment, y=RDWtotal, color=Treatment)) + 
      geom_jitter(shape=1, width=0.1) + 
      labs(x = "", y = "Total root dry weight (mg)") +
      theme_bw()+
      stat_summary(fun.data="mean_cl_boot", size=0.5)+
      scale_color_manual(values=colors)+
      theme(legend.position = "none",
            axis.text=element_text(size=10, color="black"),
            axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
            axis.text.x=element_text(angle=0))+
      scale_y_continuous(limits=c(0,300), breaks=c(0,50,100,150,200,250,300)))

(p<-ggplot(root1, aes(x=Treatment, y=invCV, color=Treatment)) + 
      geom_jitter(shape=1, width=0.1) + 
      labs(x = "", y = "Community inverse CV") +
      theme_bw()+
      stat_summary(fun.data="mean_cl_boot", size=0.5)+
      scale_color_manual(values=colors)+
      theme(legend.position = "none",
            axis.text=element_text(size=10, color="black"),
            axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
            axis.text.x=element_text(angle=0)))

#ggsave("Total_root_productivity.tiff", p3, dpi=1000, compression="lzw", width=9, height=9, units="cm")
```

### Effect sizes

#### Total root productivity

```{r, fig.width=4, fig.height=4}
es.rdw<-bootstrap_ES(response="RDWtotal", data=root1, contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.rdw$summary)

ggplot(es.rdw$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.rdw$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.rdw$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES RDW (mg)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

#### MRD biomass

```{r, fig.width=4, fig.height=4}
es.mrdb<-bootstrap_ES(response="MRD", data=root1, contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.mrdb$summary)

ggplot(es.mrdb$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.mrdb$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.mrdb$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES MRD (cm)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

### Plot root biomass distribution in soil layers

```{r Plot root biomass distribution, message=FALSE, warning=FALSE, fig.width=7.09, fig.height=5.12}

supp.labs<-c("0-10 cm", "10-20 cm", "20-30 cm", "30-40 cm", "40-50 cm", ">50 cm")
names(supp.labs)<-c("0-10 cm", "10-20 cm", "20-30 cm", "30-40 cm", "40-50 cm", "50-60 cm")
  
(p4<-ggplot(root, aes(x=Treatment, y=RDW, color=Treatment)) + 
    geom_jitter(shape=1, width=0.1) + 
    labs(x = "", y = "Total root dry weight (mg)") +
    theme_bw()+
    stat_summary(fun.data="mean_cl_boot", size=0.5)+
    scale_color_manual(values=colors)+
    theme(legend.position = "none",
          axis.text=element_text(color="black"),
          axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
          axis.text.x=element_text(size=10, angle=45, hjust = 1, vjust=1),
          axis.text.y=element_text(size=10),
          strip.text = element_text(size=10))+
    facet_wrap(~Layer, scales="free_y",
               labeller=labeller(Layer=supp.labs)))

#ggsave("FigureS3.tiff", p4, dpi=1000, compression="lzw", width=18, height=13, units="cm")
```

### Root biomass distribution in the soil

Because some layers did not contain any roots (RDW=0), we had zero-inflated continuous data ($Y \ge 0$). Data were analysed using a hurdle model combining a gamma GLMM and a Bernoulli GLMM (zero-altered gamma model or ZAG model). In a ZAG model, zeroes are generated by the binary part of the model (1 or 0). Non-zero data (that occur that roots are present in a layer) are modeled using a second (continuous) distribution (gamma in this case). The hurdle model was fitted according to Chapter 7 in Zuur and Ieno (2016) [Beginner's guide to zero-inflated models with R](http://highstat.com/index.php/beginner-s-guide-to-zero-inflated-models). Highland Statistics Ltd. 

#### Fit gamma component of hurdle model

```{r Fit gamma model, message=FALSE, warning=FALSE}
root$Depth<-root$maxDepth-5

root.pos<-subset(root, RDW>0) #Only select positive RDW values
root.pos$Rz<-factor(root.pos$Rz)

#Fit GLMM with gamma distribution
H1<-glmmPQL(fixed=RDW~Depth*Treatment, random=~1|Rz, data=root.pos, family=Gamma(link="log"))
summary(H1)

#Model validation
E1<-resid(H1, type="pearson")
F1<-fitted(H1, type="response")
plot(F1, E1, xlab="Fitted values", ylab="Residuals"); abline(h=0)
boxplot(E1~root.pos$Treatment, xlab="", ylab="Residuals")
boxplot(E1~root.pos$Depth, xlab="Depth (cm)", ylab="Residuals")

#Make predictions
pred.mm.H1 <- data.frame(Depth=rep(seq(5,55,by=0.1), 5), Treatment=rep(unique(root$Treatment), each=501))
pred.mm.H1$Pred <- predict(H1, newdata=pred.mm.H1, type="response", level=0)

#Plot GLMM fit
(p5<-ggplot(pred.mm.H1) +
    geom_jitter(data=root.pos, aes(x=Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(aes(x = Depth, y = Pred, color=Treatment), size=1) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))

#Plot GLMM fit
#One curve for each rhizobox
(p6<-ggplot(root.pos) +
    geom_jitter(aes(x=Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(data=cbind(root.pos, Pred=predict(H1, type="response")), aes(x = Depth, y = Pred, color=Treatment, group=Rz), size=1) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))
```
 
#### Fit binomial component of hurdle model

```{r Fit binomial model, message=FALSE, warning=FALSE}
root$RDW.01<-as.numeric(root$RDW > 0)
root$Rz<-factor(root$Rz)
root$sDepth<-scale(root$Depth, center=TRUE, scale=TRUE)

#Fit GLMM with a binomial distribution
H2<-glmmPQL(fixed=RDW.01~Depth*Treatment, random=~1|Rz, data=root, family=binomial(link="logit"))
summary(H2)

#Make predictions
pred.mm.H2 <- data.frame(Depth=rep(seq(5,55,by=0.1), 5),
                      Treatment=rep(unique(root$Treatment), each=501))
pred.mm.H2$Pred <- predict(H2, newdata=pred.mm.H2, type="response", level=0)

#Plot binomial fit
(p7<-ggplot(pred.mm.H2) +
    geom_jitter(data=root, aes(x=Depth, y=RDW.01, colour=Treatment), height=0.2, width=1, shape=1) +
    geom_line(aes(x = Depth, y = Pred, color=Treatment), size=1) +
    ylab("Probability that roots are present")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors))
```

#### Merge gamma and binomial components

```{r Create hurdle model, message=FALSE, warning=FALSE, fig.height=3.94, fig.width=3.94}
beta<-fixef(H1) #Extract regression coefficients of Gamma GLMM
gamma<-fixef(H2) #Extract regression coefficients of Binomial GLMM

#Create new data for predictions
newdata <- expand.grid(Depth = seq(5, 55, by = 0.1),
                       Treatment = levels(root$Treatment))

X<-model.matrix(~Depth*Treatment, data=newdata)

mu <- exp(X %*% beta)
Pi  <- exp(X %*% gamma) / (1 + exp(X %*% gamma))

#Get predictions for Hurdle model
newdata$PredZAG<-Pi * mu

#Plot hurdle model fit
(p8<-ggplot(root) +
    geom_jitter(data=root, aes(x=-Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(data=newdata, aes(x = -Depth, y = PredZAG, color=Treatment), size=0.8) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          legend.position=c(0.85,0.22),
          legend.text=element_text(size=8),
          legend.background = element_blank(),
          legend.key = element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_x_continuous(breaks=c(-5,-15,-25,-35,-45,-55))+
    scale_color_manual(values=colors)+
    coord_flip())

#ggsave("Figure4.tiff", p8, dpi=1000, compression="lzw", width=9.5, height=10, units="cm")
```

#### Create figure showing model fits for the Gamma and Binomial components of the hurdle model

```{r Create Figure S2, message=FALSE, warning=FALSE, fig.height=3.94, fig.width=7.09}
#Plot gamma component
p5<-ggplot(pred.mm.H1) +
    geom_jitter(data=root.pos, aes(x=-Depth, y=RDW, colour=Treatment), width=1, shape=1) +
    geom_line(aes(x = -Depth, y = Pred, color=Treatment), size=1) +
    ylab("Root dry weight (mg)")+
    xlab("Depth (cm)")+ 
    theme_bw()+
    theme(legend.title=element_blank(),
          legend.position=c(0.85,0.22),
          legend.text=element_text(size=9),
          legend.background = element_blank(),
          legend.key = element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors)+
    coord_flip()+
    scale_x_continuous(breaks=c(-5,-15,-25,-35,-45,-55))+
    ggtitle("Gamma component")

#Plot binomial component
p7<-ggplot(pred.mm.H2) +
    geom_jitter(data=root, aes(x=-Depth, y=RDW.01, colour=Treatment), height=0.1, width=1, shape=1) +
    geom_line(aes(x = -Depth, y = Pred, color=Treatment), size=1) +
    ylab("Probability that roots are present")+
    xlab("")+ 
    theme_bw()+
    theme(legend.position="none",
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_color_manual(values=colors)+
    coord_flip()+
    scale_x_continuous(breaks=c(-5,-15,-25,-35,-45,-55))+
    scale_y_continuous(breaks=c(0,0.5,1))+
    ggtitle("Binomial component")

(p5and7<-ggarrange(p5, p7, nrow=1, ncol=2, labels=c("(a)", "(b)"), font.label = list(size=12)))

#ggsave("FigureS2.tiff", p5and7, dpi=1000, compression="lzw", width=18, height=10, units="cm")
```

## Analysis of root distribution over time

### Load RootPainter data

```{r Load RootPainter data, message=FALSE, warning=FALSE}
root.depth <- read.csv("Z:/Poem/Rhizoboxes/Backup_RootPainter/drive_rp_sync/projects/PE_rhizotrons_2017_1/results/segmentations/results.csv")

root.length <- read.csv("Z:/Poem/Rhizoboxes/Backup_RootPainter/drive_rp_sync/projects/PE_rhizotrons_2017_1/results/analysis_RVE/features.csv")

ID <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Image_RZ.txt")
TIME <- read.delim("Z:/Poem/Rhizoboxes/0_Rz_experiment_data/Data/Date_TimePoint.txt")

filenames<-sapply(str_split(as.character(root.depth$image), pattern="-"), function(x){return(x[5])})
filenames<-as.numeric(sapply(str_split(filenames, pattern="\\."), function(x){return(x[1])}))

root.depth$Date<-sapply(str_split(as.character(root.depth$image), pattern="-"), function(x){
  return(paste(x[1], x[2], x[3], sep="-"))})

root.depth$imageID<-filenames
root.depth$Rz<-ID$RZ[match(filenames, ID$Image)]
root.depth$Time<-TIME$Day[match(root.depth$Date, TIME$Date)]

root.depth<-root.depth[order(root.depth$Rz, root.depth$Time),]

#Add treatment column
root.depth$Treatment<-root1$Treatment[match(root.depth$Rz, root1$Rz)]

#Add root_length column in root.depth
root.depth$root_length<-root.length$Total.Root.Length.mm[match(root.depth$image, root.length$File.Name)]
root.depth$root_length[is.na(root.depth$root_length)==T]<-0 #Replace NA values by zero
root.depth$root_length<-root.depth$root_length/1000 #Express results in meters

#Set Sync2 to zero before the second sowing
root.depth$mrd[which(root.depth$Treatment=="Sync2" & root.depth$Time<=9)]<-0
root.depth$maxrd[which(root.depth$Treatment=="Sync2" & root.depth$Time<=9)]<-0
root.depth$root_length[which(root.depth$Treatment=="Sync2" & root.depth$Time<=9)]<-0

#Remove rhizoboxes 35, 70, 15, 27, 137
root.depth<-root.depth[-which(root.depth$Rz==35|root.depth$Rz==70|root.depth$Rz==15|root.depth$Rz==27|root.depth$Rz==137),]

root.depth$Treatment<-factor(root.depth$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))

#No root in image "2017-08-16-RZ-0001.png", so set mrd and maxrd to zero
root.depth$mrd[root.depth$image=="2017-08-16-RZ-0001.png"]<-0
root.depth$maxrd[root.depth$image=="2017-08-16-RZ-0001.png"]<-0
root.depth$root_length[root.depth$image=="2017-08-16-RZ-0001.png"]<-0
```

### Plot mean rooting depth over time

```{r Calculate MRD, message=FALSE, warning=FALSE}
#Create empty dataset for summary statistics
stat.mrd<-data.frame(Treatment=rep(NA, length(unique(root.depth$Rz))),
                     Time=rep(NA, length(unique(root.depth$Rz))),
                     meanValue=rep(NA, length(unique(root.depth$Rz))),
                     lower=rep(NA, length(unique(root.depth$Rz))),
                     upper=rep(NA, length(unique(root.depth$Rz))))
 
k<-0
 
for (i in unique(root.depth$Treatment)){
   
  for (j in unique(root.depth$Time)){
     
    k<-k+1
    sub<-root.depth[root.depth$Treatment==i & root.depth$Time==j,]
    #Calculate compatibility interval (bootstrap resampling)
    set.seed(1)
    a<-smean.cl.boot(sub$mrd, B=10000)
    stat.mrd[k,]<-c(i, j, as.numeric(a))}}
 
stat.mrd[,2]<-as.numeric(stat.mrd[,2])
stat.mrd[,3]<-as.numeric(stat.mrd[,3])
stat.mrd[,4]<-as.numeric(stat.mrd[,4])
stat.mrd[,5]<-as.numeric(stat.mrd[,5])
 
stat.mrd$Treatment<-factor(stat.mrd$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))

#Plot mean rooting depth over time
(p9<-ggplot(stat.mrd, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MRD (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(stat.mrd$Time))+
  scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1)))

#Plot mean rooting depth over time
#One line per rhizobox
(p10<-ggplot(root.depth, aes(x=Time, y=mrd, color=Treatment, group=Rz))+
  geom_line()+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MRD (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors))

#Plot mean rooting depth at last observation date

(p11<-ggplot(root.depth[root.depth$Time==44,], aes(x=Treatment, y=mrd, colour=Treatment)) +
    geom_jitter(width=0.1, shape=1) + 
    stat_summary(fun.data="mean_cl_boot")+
    ylab("MRD (cm)")+
    xlab("")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10))))

model<-glm(-mrd~Treatment, root.depth[root.depth$Time==44,], family=Gamma(link="log"))

summary(model)
summary(emmeans(model, "Treatment", contr="tukey"), type="response")
summary(emmeans(model, "Treatment", contr="tukey"))
```

### Plot maximum rooting depth over time

```{r Calculate MaxRD, message=FALSE, warning=FALSE}
#Create empty dataset for summary statistics
stat.maxrd<-data.frame(Treatment=rep(NA, length(unique(root.depth$Rz))),
                       Time=rep(NA, length(unique(root.depth$Rz))),
                       meanValue=rep(NA, length(unique(root.depth$Rz))),
                       lower=rep(NA, length(unique(root.depth$Rz))),
                       upper=rep(NA, length(unique(root.depth$Rz))))
 
k<-0
 
for (i in unique(root.depth$Treatment)){
   
  for (j in unique(root.depth$Time)){
     
    k<-k+1
    sub<-root.depth[root.depth$Treatment==i & root.depth$Time==j,]
    #Calculate compatibility interval (bootstrap resampling)
    set.seed(1)
    a<-smean.cl.boot(sub$maxrd, B=10000)
    stat.maxrd[k,]<-c(i, j, as.numeric(a))}}
 
stat.maxrd[,2]<-as.numeric(stat.maxrd[,2])
stat.maxrd[,3]<-as.numeric(stat.maxrd[,3])
stat.maxrd[,4]<-as.numeric(stat.maxrd[,4])
stat.maxrd[,5]<-as.numeric(stat.maxrd[,5])
 
stat.maxrd$Treatment<-factor(stat.maxrd$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))

#Plot max rooting depth over time
(p9a<-ggplot(stat.maxrd, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MaxRD (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(stat.maxrd$Time))+
  scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1)))

#Plot max rooting depth over time
#One line per rhizobox
(p10a<-ggplot(root.depth, aes(x=Time, y=maxrd, color=Treatment, group=Rz))+
  geom_line()+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MaxRD (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors))

#Plot max rooting depth at last observation date

(p11a<-ggplot(root.depth[root.depth$Time==44,], aes(x=Treatment, y=maxrd, colour=Treatment)) +
    geom_jitter(width=0.1, shape=1) + 
    stat_summary(fun.data="mean_cl_boot")+
    ylab("MaxRD (cm)")+
    xlab("")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10))))

model<-glm(-maxrd~Treatment, root.depth[root.depth$Time==44,], family=Gamma(link="log"))

summary(model)
summary(emmeans(model, "Treatment", contr="tukey"), type="response")
summary(emmeans(model, "Treatment", contr="tukey"))
```

### Plot total root length over time

```{r Calculate TRL, message=FALSE, warning=FALSE}
#Create empty dataset for summary statistics
stat.root_length<-data.frame(Treatment=rep(NA, length(unique(root.depth$Rz))),
                       Time=rep(NA, length(unique(root.depth$Rz))),
                       meanValue=rep(NA, length(unique(root.depth$Rz))),
                       lower=rep(NA, length(unique(root.depth$Rz))),
                       upper=rep(NA, length(unique(root.depth$Rz))))
 
k<-0
 
for (i in unique(root.depth$Treatment)){
   
  for (j in unique(root.depth$Time)){
     
    k<-k+1
    sub<-root.depth[root.depth$Treatment==i & root.depth$Time==j,]
    #Calculate compatibility interval (bootstrap resampling)
    set.seed(1)
    a<-smean.cl.boot(sub$root_length, B=10000)
    stat.root_length[k,]<-c(i, j, as.numeric(a))}}
 
stat.root_length[,2]<-as.numeric(stat.root_length[,2])
stat.root_length[,3]<-as.numeric(stat.root_length[,3])
stat.root_length[,4]<-as.numeric(stat.root_length[,4])
stat.root_length[,5]<-as.numeric(stat.root_length[,5])
 
stat.root_length$Treatment<-factor(stat.root_length$Treatment, levels=c("Sync1", "F-first", "G-first", "L-first", "Sync2"))

#Plot total root length over time
(p9b<-ggplot(stat.root_length, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("Total root length (m)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(stat.root_length$Time))+
  scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1)))

#Plot total root length over time
#One line per rhizobox
(p10b<-ggplot(root.depth, aes(x=Time, y=root_length, color=Treatment, group=Rz))+
  geom_line()+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("Total root length (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors))

#Plot total root length at last observation date

(p11b<-ggplot(root.depth[root.depth$Time==44,], aes(x=Treatment, y=root_length, colour=Treatment)) +
    geom_jitter(width=0.1, shape=1) + 
    stat_summary(fun.data="mean_cl_boot")+
    ylab("Total root length (m)")+
    xlab("")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10))))

#Calculate and plot effect sizes for total root length
es.trl<-bootstrap_ES(response="root_length", data=root.depth[root.depth$Time==44,], contr=c("F-first/G-first", "F-first/L-first", "G-first/L-first"), plot=FALSE)
knitr::kable(es.trl$summary)

ggplot(es.trl$bootstrap)+
  geom_hline(yintercept = 0)+
  stat_ydensity(aes(x=Contrast, y=ES), fill="grey50", color=adjustcolor("white", alpha.f = 0.1), alpha=0.3)+
  geom_segment(data=es.trl$summary, aes(x=Contrast, xend=Contrast, y=lowCI, yend=highCI), color="black", size=0.5)+
  geom_point(data=es.trl$summary, aes(x=Contrast, y=ES), shape=16, color="black")+
  theme_bw()+
  xlab("")+
  ylab("ES TRL (m)")+
  theme(axis.title=element_text(color="black", face="plain"),
  axis.title.y=element_text(margin = margin(t = 0, r = 5, b = 0, l = 0)),
  axis.title.x=element_blank(),
  axis.text=element_text(color="black"))
```

### Create combined plot for aboveground and belowground productivity and total root length

```{r Create figure 2, message=FALSE, warning=FALSE, fig.height=6.3, fig.width=7.09}
p2<-ggplot(shoot, aes(x=Treatment, y=Total, color=Treatment)) + 
    geom_jitter(shape=1, width=0.15) + 
    labs(x = "", y = "Total shoot dry weight (mg)") +
    theme_bw()+
    stat_summary(fun.data="mean_cl_boot", size=0.5)+
    geom_signif(y_position=740, xmin=2, xmax=3, annotations="", linetype=2, tip_length = 0.02,
                color="black")+
    geom_signif(y_position=740, xmin=3, xmax=4, annotations="", linetype=2, tip_length = 0.02,
                color="black")+
    scale_color_manual(values=colors)+
    theme(legend.position = "none",
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(t=0,b=0,r=5,l=0)),
          axis.text.x=element_text(angle=0))

p3<-ggplot(root1, aes(x=Treatment, y=RDWtotal, color=Treatment)) + 
      geom_jitter(shape=1, width=0.15) + 
      labs(x = "", y = "Total root dry weight (mg)") +
      theme_bw()+
      stat_summary(fun.data="mean_cl_boot", size=0.5)+
      geom_signif(y_position=180, xmin=2, xmax=3, annotations="", linetype=2, tip_length = 0.02,
                color="black")+
      geom_signif(y_position=180, xmin=3, xmax=4, annotations="", linetype=2, tip_length = 0.02,
                color="black")+
      scale_color_manual(values=colors)+
      theme(legend.position = "none",
            axis.text=element_text(size=10, color="black"),
            axis.title.y = element_text(margin=margin(t=0,b=0,r=10,l=0)),
            axis.text.x=element_text(angle=0))+
      scale_y_continuous(limits=c(0,300), breaks=c(0,50,100,150,200,250,300))

p9b<-ggplot(stat.root_length, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        legend.position=c(0.36,0.79),
        legend.text=element_text(size=9),
        legend.key.size = unit(1, "line"),
        legend.background = element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("Total root length (m)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1))+
  scale_x_continuous(breaks=unique(stat.root_length$Time),
                     labels=c(2,4,"",9,"","",16,"","",23,"","",30,"","",37, "","",44))

p11b<-ggplot(root.depth[root.depth$Time==44,], aes(x=Treatment, y=root_length, colour=Treatment)) +
    geom_jitter(width=0.15, shape=1) + 
    stat_summary(fun.data="mean_cl_boot")+
    ylab("Total root length, day 44 (m)")+
    xlab("")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    theme(legend.position="none",
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    geom_signif(y_position=15.5, xmin=3, xmax=4, annotations="", linetype=2, tip_length = 0.02,
                color="black")+
    geom_signif(y_position=17.5, xmin=2, xmax=3, annotations="*", linetype=1, tip_length = 0.02,
                color="black", textsize = 5)+
    geom_signif(y_position=20, xmin=2, xmax=4, annotations="*", linetype=1, tip_length = 0.02,
                color="black", textsize = 5)

(figure2<-ggarrange(p2, p3, p9b, p11b, nrow=2, ncol=2, labels=c("(a)", "(b)", "(c)", "(d)"), font.label=list(size=14), hjust=-0.2, align="hv"))

#ggsave("Figure2.tiff", figure2, dpi=1000, compression="lzw", width=18, height=16, units="cm")
```

### Fit linear mixed effect model for MRD

```{r Fit lmer to mean rooting depth data, message=FALSE, warning=FALSE}
#Calculate number of days after first sowing (dafs)
root.depth$dafs<-root.depth$Time
root.depth$dafs[root.depth$Treatment=="Sync2"]<-root.depth$dafs[root.depth$Treatment=="Sync2"]-10
root.depth1<-root.depth[root.depth$dafs>=0,]

#Fit linear mixed-effect model(random slope and random intercept)
model<-lmer(mrd~Treatment*dafs+(dafs|Rz), root.depth1, 
            control = lmerControl(optimizer ="Nelder_Mead"))

summary(model)
plot(model)

#Plot model fit
#One line per rhizobox
(p12<-ggplot(root.depth1, aes(x=dafs, y=mrd, color=Treatment, group=Rz))+
  geom_line(data = cbind(root.depth1, pred = predict(model)), aes(y = pred), size = 1)+
  geom_point(shape=16)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MRD (cm)")+
  xlab("Days after sowing")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(limits=c(1, 44)))

pred.mm <- ggpredict(model, terms = c("dafs [1,2,4,6,7,8,9,11,13,14,15,16,18,20,21,22,23,25,27,28,29,30,32,34,35,37,39,42,44]", "Treatment"))

data.mm.mrd<-data.frame(dafs=pred.mm$x,
                        Pred=pred.mm$predicted,
                        Treatment=pred.mm$group,
                        CIlow=pred.mm$conf.low,
                        CIhigh=pred.mm$conf.high)

data.mm.mrd<-data.mm.mrd[-which(data.mm.mrd$Treatment=="Sync2" & data.mm.mrd$dafs>34),]
data.mm.mrd<-data.mm.mrd[-which(data.mm.mrd$Treatment!="Sync2" & data.mm.mrd$dafs<2),]

#Plot model fit
(p13<-ggplot(data.mm.mrd) + 
    geom_line(aes(x = dafs, y = Pred, group=Treatment, color=Treatment), size=1) +   
    geom_ribbon(aes(x = dafs, ymin = CIlow, ymax = CIhigh, fill=Treatment, alpha=Treatment)) +
    ylab("MRD (cm)")+
    xlab("Days after sowing")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    scale_fill_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1)))
```

### Fit linear mixed effect model for MaxRD

```{r Fit lmer to maximum rooting depth data, message=FALSE, warning=FALSE}
#For Sync1, L-first and F-first, only consider time points before reaching the bottom of the rhizoboxes (remove time points 35 to 44 for Sync1 and L-first, remove time points 42 and 44 for F-first)
root.depth2<-root.depth1
root.depth2<-rbind(root.depth2[root.depth2$Time<=32 & 
                           (root.depth2$Treatment=="Sync1"|
                              root.depth2$Treatment=="L-first"),],
                   root.depth2[root.depth2$Time<=39 & root.depth2$Treatment=="F-first",],
                   root.depth2[root.depth2$Treatment=="Sync2"|
                              root.depth2$Treatment=="G-first",])

#Fit linear mixed-effect model(random slope and random intercept)
model<-lmer(maxrd~Treatment*dafs+(dafs|Rz), root.depth2, 
            control = lmerControl(optimizer ="Nelder_Mead"))

summary(model)
plot(model)

#Plot model fit
#One line per rhizobox
(p12a<-ggplot(root.depth2, aes(x=dafs, y=maxrd, color=Treatment, group=Rz))+
  geom_line(data = cbind(root.depth2, pred = predict(model)), aes(y = pred), size = 1)+
  geom_point(shape=16)+
  theme_bw()+
  theme(legend.title=element_blank(),
        axis.text=element_text(size=10, color="black"),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)))+
  ylab("MaxRD (cm)")+
  xlab("Days after sowing")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(limits=c(1, 44)))

pred.mm <- ggpredict(model, terms = c("dafs [1,2,4,6,7,8,9,11,13,14,15,16,18,20,21,22,23,25,27,28,29,30,32,34,35,37,39,42,44]", "Treatment"))

data.mm.maxrd<-data.frame(dafs=pred.mm$x,
                          Pred=pred.mm$predicted,
                          Treatment=pred.mm$group,
                          CIlow=pred.mm$conf.low,
                          CIhigh=pred.mm$conf.high)

data.mm.maxrd<-data.mm.maxrd[-which(data.mm.maxrd$Treatment=="Sync2" & data.mm.maxrd$dafs>34),]
data.mm.maxrd<-data.mm.maxrd[-which(data.mm.maxrd$Treatment!="Sync2" & data.mm.maxrd$dafs<2),]
data.mm.maxrd<-data.mm.maxrd[-which(data.mm.maxrd$Treatment=="Sync1" & data.mm.maxrd$dafs>32),]
data.mm.maxrd<-data.mm.maxrd[-which(data.mm.maxrd$Treatment=="L-first" & data.mm.maxrd$dafs>32),]
data.mm.maxrd<-data.mm.maxrd[-which(data.mm.maxrd$Treatment=="F-first" & data.mm.maxrd$dafs>39),]

#Plot model fit
(p13a<-ggplot(data.mm.maxrd) + 
    geom_line(aes(x = dafs, y = Pred, group=Treatment, color=Treatment), size=1) +   
    geom_ribbon(aes(x = dafs, ymin = CIlow, ymax = CIhigh, fill=Treatment, alpha=Treatment)) +
    ylab("MaxRD (cm)")+
    xlab("Days after sowing")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    scale_fill_manual(values=colors)+
    theme(legend.title=element_blank(),
          axis.text=element_text(size=10, color="black"),
          axis.title.y = element_text(margin=margin(r=10)),
          axis.title.x = element_text(margin=margin(t=10)))+
    scale_alpha_manual(values=c(0.3,0.2,0.2,0.2,0.1)))
```

### Correlation between MRD biomass and MRD images

```{r Correlation MRD, fig.height=3.94, fig.width=3.94, message=FALSE, warning=FALSE}
root1<-root1[order(root1$Rz),] #Contain MRD values calculated from root biomass
root1$MRDrp<-root.depth[root.depth$Time==44,"mrd"]

corrMRD<-data.frame(MRDb=root1$MRD,
                    MRDi=root1$MRDrp,
                    Treatment=root1$Treatment)

cor.test(corrMRD$MRDb, -corrMRD$MRDi)

text.size<-8
title.size<-9

(p14<-ggplot(corrMRD, aes(x=-MRDi, y=MRDb, colour=Treatment))+
      geom_smooth(method="lm", colour="black", size=1)+
      geom_point(shape=16)+
      theme_bw()+
      theme(legend.title=element_blank(),
            legend.position=c(0.88,0.16),
            legend.text=element_text(size=8),
            legend.key.size = unit(0.8, "line"),
            legend.background = element_blank(),
            legend.key = element_blank(),
            axis.text=element_text(size=text.size, color="black"),
            axis.title=element_text(size=title.size),
            axis.title.y = element_text(margin=margin(r=5)),
            axis.title.x = element_text(margin=margin(t=5)))+
      scale_color_manual(values=colors)+
      xlab("MRD - image analysis (cm)")+
      ylab("MRD - biomass (cm)"))

#ggsave("FigureS1.tiff", p14, dpi=1000, compression="lzw", width=10, height=10, units="cm")
```

### Create figure MRD and MaxRD over time

```{r Figure MRD and MaxRD over time, message=FALSE, warning=FALSE, fig.height=5.12, fig.width=6.3}
text.size<-8
title.size<-9

p9<-ggplot(stat.mrd, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16, size=1)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.title=element_blank(),
        legend.position=c(0.87,0.78),
        legend.text=element_text(size=8),
        legend.key.size = unit(0.9, "line"),
        legend.background = element_blank(),
        axis.text=element_text(size=text.size, color="black"),
        axis.title=element_text(size=title.size),
        axis.title.y = element_text(margin=margin(r=5)),
        axis.title.x = element_blank(),
        panel.grid.minor = element_blank())+
  ylab("Mean rooting depth (cm)")+
  xlab("")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(root.depth$Time),
                     labels=c(2,4,"",9,"","",16,"","",23,"","",30,"","",37, "","",44))+
  scale_alpha_manual(values=c(0.2,0.2,0.2,0.2,0.1))

p13<-ggplot(data.mm.mrd) + 
    geom_line(aes(x = dafs, y = Pred, group=Treatment, color=Treatment), size=0.6) +
    geom_ribbon(aes(x = dafs, ymin = CIlow, ymax = CIhigh, fill=Treatment, alpha=Treatment), color=NA) +
    ylab("")+
    xlab("")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    scale_fill_manual(values=colors)+
    theme(legend.position="none",
          axis.text=element_text(size=text.size, color="black"),
          axis.title=element_text(size=title.size),
          axis.title.y = element_text(margin=margin(r=5)),
          axis.title.x = element_blank(),
          panel.grid.minor = element_blank())+
    scale_alpha_manual(values=c(0.2,0.2,0.2,0.2,0.1))+
    scale_x_continuous(breaks=seq(0,44, by=4),
                     labels=seq(0,44, by=4))+
    scale_y_continuous(breaks=c(0,-4,-8,-12,-16), labels=c(0,-4,-8,-12,-16), limits=c(-16,0))

p9a<-ggplot(stat.maxrd, aes(x=Time, y=meanValue, color=Treatment))+
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Treatment, alpha=Treatment), color=NA)+
  geom_line()+
  geom_point(shape=16, size=1)+
  geom_vline(xintercept=10, linetype=3)+
  theme_bw()+
  theme(legend.position = "none",
        axis.text=element_text(size=text.size, color="black"),
        axis.title=element_text(size=title.size),
        axis.title.y = element_text(margin=margin(r=5)),
        axis.title.x = element_text(margin=margin(t=5)),
        panel.grid.minor = element_blank(),
        plot.title = element_blank())+
  ylab("Maximum rooting depth (cm)")+
  xlab("Time (days)")+
  scale_color_manual(values=colors)+
  scale_fill_manual(values=colors)+
  scale_x_continuous(breaks=unique(root.depth$Time),
                     labels=c(2,4,"",9,"","",16,"","",23,"","",30,"","",37, "","",44))+
  scale_alpha_manual(values=c(0.2,0.2,0.2,0.2,0.1))+
  scale_y_continuous(breaks=c(0,-10,-20,-30,-40,-50,-60), labels=c(0,-10,-20,-30,-40,-50,-60), limits=c(-60,0))

p13a<-ggplot(data.mm.maxrd) + 
    geom_line(aes(x = dafs, y = Pred, group=Treatment, color=Treatment), size=0.6) +
    geom_ribbon(aes(x = dafs, ymin = CIlow, ymax = CIhigh, fill=Treatment, alpha=Treatment), color=NA) +
    ylab("")+
    xlab("Number of days after sowing")+ 
    theme_bw()+
    scale_color_manual(values=colors)+
    scale_fill_manual(values=colors)+
    theme(legend.position="none",
          axis.text=element_text(size=text.size, color="black"),
          axis.title=element_text(size=title.size),
          axis.title.y = element_text(margin=margin(r=5)),
          axis.title.x = element_text(margin=margin(t=5)),
          panel.grid.minor = element_blank(),
          plot.title = element_blank())+
    scale_alpha_manual(values=c(0.2,0.2,0.2,0.2,0.1))+
    scale_x_continuous(breaks=seq(0,44, by=4),
                     labels=seq(0,44, by=4))+
    scale_y_continuous(breaks=c(0,-10,-20,-30,-40,-50,-60), labels=c(0,-10,-20,-30,-40,-50,-60))

(p15<-ggarrange(p9, p13, p9a, p13a, nrow=2, ncol=2, labels=c("(a)", "(b)", "(c)", "(d)"), font.label = list(size=11), 
               heights = c(0.93,1)))

#ggsave("Figure5.tiff", p15, dpi=1000, compression="lzw", width=16, height=13, units="cm")
```